Formulating and Validating RDF Constraints Generically

### 2015-01-02 paper/paper 2/paper 2.tex
automatic validation

@@

\keywords{..}
\end{abstract}
%

@@

\begin{itemize}
	\item requirements database
	\item relationship between requirements and constraints
	\item constraint languages
\end{itemize}

@@

In Section 1.3 we introduced three forms of RBox axioms: role inclusions, role equiv-
alences and role disjointness. OWL provides a variety of others, namely role transi-
tivity, symmetry, asymmetry, reflexivity and irreflexivity. These are sometimes consid-
ered as basic axiom types in DLs as well, using some suggestive notation such as
Trans( ancestorOf ) to express that the role ancestorOf is transitive. However, such axioms
are just syntactic sugar; all role characteristics can be expressed using the features
of DLs that we have already introduced.

\begin{itemize}
  \item domain and range
	\item Primary Key Properties
	\item (ir)reflexive object properties
	\item ( Exact Qualified Cardinality Restrictions on Properties )
\end{itemize}

@@

\section{Transformations between Specific Constraints}

\section{Implementation}

>>>>

%\usepackage{listings}% http://ctan.org/pkg/listings
%\lstset{
  %numbers=left,numbersep=2mm,frame=single,mathescape,fontsize=\scriptsize
%}

@@

\keywords{RDF Validation, RDF Constraints, RDF Validation Requirements, Linked Data, Semantic Web}
\end{abstract}
%

@@

For many RDF applications, the formulation of constraints and the automatic validation of data according to these constraints is a much sought-after feature. 
In 2013, the W3C invited experts from industry, government, and academia to the RDF Validation Workshop\footnote{\url{http://www.w3.org/2012/12/rdf-val/}}, 
where first use cases have been presented and discussed. 
Two working groups (WGs),  that follow up on this workshop and address RDF constraint formulation and validation, are established in 2014: 
the W3C RDF Data Shapes WG\footnote{\url{http://www.w3.org/2014/rds/charter}} and the DCMI RDF Application Profiles WG\footnote{\url{http://wiki.dublincore.org/index.php/RDF-Application-Profiles}}.

Bosch and Eckert \cite{BoschEckert2014} initiated a database on RDF validation requirements to collect case studies, use cases and especially requirements regarding RDF validation collaboratively and in a comprehensive and structured way. 
%\er{Just for reminding ourselves:  being careful using what is a use-case and what is a case-study later in the text.(http://www.w3.org/2001/sw/sweo/public/UseCases/)}
%\tb{agreed, will take care}
This database is continuously updated and openly available at \url{http://purl.org/net/rdf-validation}. It is used to evaluate and compare several existing solutions for RDF constraint formulation and validation. 
The requirements are classified to provide a high-level view on different solutions and to facilitate a better understanding of the problem domain \cite{BoschEckert2014}.  
Each requirement to formulate RDF constraints is directly mapped to an RDF constraint.

There are multiple constraint languages with different syntaxes and semantics which can be used to express RDF constraints, e.g., \emph{at-least} restrictions.
The most promising ones on being the standard, are
the SPARQL Inferencing Notation (SPIN)\footnote{\url{http://spinrdf.org/}}, 
the Web Ontology Language (OWL 2)\footnote{\url{http://www.w3.org/TR/owl2-syntax/}}, 
Shape Expressions (ShEx)\footnote{\url{http://www.w3.org/Submission/shex-primer/}}, 
Resource Shapes (ReSh)\footnote{\url{http://www.w3.org/Submission/shapes/}}, 
and Description Set Profiles (DSP)\footnote{\url{http://dublincore.org/documents/2008/03/31/dc-dsp/}}.
However, there is no single standard general-purpose constraint language.
The choice depends on the individual use case.

\tb{copied from first paper - need to be modified}

@@

There three forms of RBox axioms: role inclusions, role equivalences and role disjointness. 
OWL provides a variety of others, namely role transitivity, symmetry, asymmetry, reflexivity and irreflexivity. 
These are sometimes considered as basic axiom types in DLs as well, using some suggestive notation such as
\ms{Trans(ancestorOf)} to express that the role ancestorOf is transitive. 
However, such axioms are just syntactic sugar; 
all role characteristics can be expressed using the basic features of DLs.
\ms{Irreflexive Object Properties}\footnote{corresponds to the requirement \ms{R-60-IRREFLEXIVE-OBJECT-PROPERTIES}}
restricts that an object property is irreflexive if it is never locally reflexive, i.e. no individual is connected by the object property to itself \cite{Kroetzsch2012}.
Irreflexivity constraints, stating that individuals cannot be married to themselves, can be represented in DL:

\begin{DL}
$\top$ $\sqsubseteq$ $\neg$ $\exists  marriedTo . Self$
\end{DL}

When mapped to the RDF-CO, the complex constraint aggregates three simple constraints (one property and two class constraints):

\begin{gcotable}
property & $\exists$ marriedTo . Self & marriedTo & - & Self & $\exists$ & - \\
class & $\neg$ $\exists$ marriedTo . Self & - & - & $\exists$ marriedTo . Self & $\neg$ & - \\
class & $\top$ & - & - & $\top$, $\neg$ $\exists$ marriedTo . Self & $\sqsubseteq$ & - \\
\end{gcotable}

When using the RBox axiom \ms{role irreflexivity} as syntactic sugar, 
the complex constraint transforms into a simple property constraint with exactly the same semantics:

\begin{gcotable}
property & $\top$ & marriedTo & - & - & irreflexive & - \\
\end{gcotable}

The \ms{Primary Key Properties} constraint is often useful to declare a given (datatype) property as the "primary key" of a class, so that the system can enforce uniqueness. 
Starfleet officers are uniquely identified by their command authorization code (e.g. to activate and cancel auto-destruct sequences).
It means that the property \ms{commandAuthorizationCode} is inverse functional - mapped to DL and RDF-CO as follows:
\begin{DL}
$(\ms{funct } commandAuthorizationCode\sp{\overline{\ }})$
\end{DL}
\begin{gcotable}
property & $\top$ & commandAuthorizationCode$^{-}$ & commandAuthorizationCode & - & inverse & - \\
property & $\top$ & commandAuthorizationCode$^{-}$ & - & - & functional & - \\
\end{gcotable}

Keys are even more general, i.e., a generalization of inverse functional properties\footnote{\url{http://www.w3.org/TR/2012/REC-owl2-rdf-based-semantics-20121211/#Semantic_Conditions_for_Keys}}.
A key can be a datatype property, an object property, or a chain of properties.
For this generalization purposes, as there are different sorts of key, and as keys can lead to undecidability, 
DL is extended with \ms{key boxes} and a special \ms{keyfor} construct\cite{Lutz2005}.
This leads to the following DL and RDF-CO mappings (only one simple property constraint) for the example:

\begin{DL}
commandAuthorizationCode \ms{keyfor} StarfleetOfficer
\end{DL}

\begin{gcotable}
property & StarFleetOfficer & commandAuthorizationCode & - & - & keyfor & - \\
\end{gcotable}

%\begin{itemize}
  %\item domain and range
	%\item Primary Key Properties
	%\item ( Exact Qualified Cardinality Restrictions on Properties )
%\end{itemize}

@@

\section{Transformations and Automatic Validation of Specific Constraints}
%\section{Transformations between Specific Constraints (Expressed by Different Constraint Languages)}

RDF data providers ensure that their data conforms to constraints which are part of the contract between the providers and the consumers of that data.
As there is no standard way, these constraints may be represented by multiple constraint languages (there are even more than the five promising ones)\footnote{i.a. Bibframe, DQTP, Pellet ICV, RDFUnit, SPARQL, Stardog ICV}.
There is an obvious need to transform a specific constraint (\ms{$sc_{\alpha}$}) (expressed by any constraint language \ms{$\alpha$}) into a specific constraint (\ms{$sc_{\beta}$}) (expressed by any constraint language \ms{$\beta$}) - when they have the same semantics.
We use the RDF-CO as intermediate language for these transformations.
If we define mappings ($m(sc_{\alpha}), m(sc_{\beta})$) from these two specific constraints to the corresponding generic constraint (\ms{gc}),
we can convert them automatically:

\begin{DL}
$ gc = m(sc_{\alpha}) = m(sc_{\beta}) $
\end{DL}

We do not need to define mappings for each pair of semantically equivalent specific constraint.
Let's assume we are able to express a constraint by 5/6/7/8/9/10 constraint languages.
Without the intermediate generic constraint, we would have to define one mapping for each pair of specific constraints expressed by a specific constraint language
- that are \ms{\( {n \choose 2} \)} mappings for each constraint (example: 10/15/21/28/36/45).
Without an intermediate generic constraint, we only need to define \ms{n} mappings from the specific constraints to the generic constraint (example: 5/6/7/8/9/10 ).

%(n(n+1))/2

%\begin{DL}
%\(\sum \limits_{i=1}^n i = \frac{n(n+1)}{2}\)
%\(\sum \limits_{i=1}^n i\)
%\(\Binom{n}{2} \)
%\( {n \choose 2} \)
%\end{DL}

\ms{Allowed Values}\footnote{correspond to the requirements \ms{R-30-ALLOWED-VALUES-FOR-RDF-OBJECTS} and \ms{R-37-ALLOWED-VALUES-FOR-RDF-LITERALS}}
is a common requirement to narrow down the value space of a property by an exhaustive enumeration of the valid values (both literals or resources). 
This is often rendered in drop down boxes or radio buttons in user interfaces.
Starfleet captains, e.g. can only be overruled by commodores and admirals\footnote{simplified}.
This complex constraint is composed of two simple constraints: \ms{universal quantification} and \ms{allowed values} - which can be represented in DL:

\begin{DL}
StarfleetCaptain $\equiv$ $\forall$ overruledBy . \{Commodore\} $\sqcup$ \{Admiral\}
\end{DL}

First, we map these DL statements to two simple constraints corresponding to the RDF-CO (we define the class \ms{CaptainOverruler} by simply enumerating its instances):

\begin{gcotable}
property & StarfleetCaptain & overruledBy & - & CaptainOverruler & $\forall$ & - \\
class & CaptainOverruler & - & - & \{Commodore\},\{Admiral\} & $\sqcup$ & - \\
\end{gcotable}

The next step is to define one mapping for each specific constraint to the generic constraint.
The specific constraints (\ms{$sc_{OWL 2}$}, \ms{$sc_{ReSh}$}, and \ms{$sc_{ShEx}$}) expressed by OWL 2, ReSh, and ShEx look like this:

\begin{ex}
# OWL 2:
Enterprise rdfs:subClassOf [ a owl:Restriction .
    owl:onProperty commandedBy .
    owl:allValuesFrom [ a owl:Class ;
        owl:oneOf ( Archer Kirk Picard ) ] ] .
# ReSh:
Enterprise a rs:ResourceShape ; rs:property [
    rs:name "commandedBy" ; rs:propertyDefinition commandedBy ;
    rs:allowedValue Archer , Kirk , Picard ;
    rs:occurs rs:Exactly-one ; ] .
# ShEx:
Enterprise {
    commandedBy (Archer Kirk Picard) }
\end{ex}

%\begin{ex}
%# DSP:
%[ a dsp:DescriptionTemplate ;
    %dsp:resourceClass StarfleetCaptain ; 
    %dsp:statementTemplate [ a dsp:NonLiteralStatementTemplate ; 
        %dsp:property overruledBy ; 
        %dsp:nonLiteralConstraint [ a dsp:NonLiteralConstraint ;
            %dsp:valueClass Commodore , Admiral ] ] ] .
%# OWL 2:
%StarfleetCaptain rdfs:subClassOf [ a owl:Restriction .
    %owl:onProperty overruledBy .
    %owl:allValuesFrom [ a owl:Class ;
        %owl:oneOf ( Commodore Admiral ) ] ] .
%# ReSh:
%%StarfleetCaptain a rs:ResourceShape ;
    %%rs:property [
        %%rs:name "overruledBy" ; rs:propertyDefinition overruledBy ;
        %%rs:valueShape Commodore ; rs:occurs rs:Zero-or-many ; ] ;
    %%rs:property [
        %%rs:name "overruledBy" ; rs:propertyDefinition overruledBy ;
        %%rs:valueShape Admiral ; rs:occurs rs:Zero-or-many ; ] ;
%# ShEx:
%:StarfleetCaptain {
    %:overruledBy @:Commodore* , :overruledBy @:Admiral* }
%\end{ex}

\section{Automatic Validation of Specific and Generic Constraints}

We can provide an automatic validation of all specific constraints (expressed by any constraint language) by defining a mapping to SPIN.
That's what we have done for all OWL 2 and DSP constructs (see section \ref{sec:implementation}).
%We do not have to define a mapping to SPIN for each specific constraint. 
%when we define a mapping to SPIN for each generic constraint.
But, if we define a mapping from a specific constraint to the corresponding generic constraint and if we define a mapping from the generic constraint to SPIN,
we do not have to define a mapping to SPIN for that specific constraint - we can perform an automatic validation for the generic and for the specific constraint as well.
As a consequence, we only have to define one SPIN mapping for each constraint corresponding to a requirement within the RDF validation requirements database.
19 of 74 RDF constraints can be expressed by at least four constraint languages \cite{BoschNolleAcarEckert2015}.
This means, that there must be 76 (four for each constraint) implementations for an automatic validation of these constraints - instead of only 19 (one for each constraint).

Within our SPIN validation environment, we define a SPIN construct template for each generic constraint\footnote{for details about the SPIN validation environment see \cite{BoschEckert2014-2}}.
The SPIN construct template contains a SPARQL CONSTRUCT query which generates constraint violation triples (\ms{spin:ConstraintViolation}) indicating the subject (\ms{spin:violationRoot}) and the properties (\ms{spin:violationPath}), causing the constraint violation, and the reason why the constraint violation has been raised (\ms{rdfs:label}).
The SPIN construct template creates the constraint violation triples if all triple patterns within the SPARQL WHERE clause match.
The SPARQL CONSTRUCT query for the universal quantification is shown below:

\begin{ex}
CONSTRUCT {
    [   a spin:ConstraintViolation ; spin:violationRoot ?subject ;
        rdfs:label ?violationMessage ; spin:violationPath ?p1 . }
WHERE {	  
    [   a gclo:PropertyConstraint ;
        gclo:contextClass ?cc ;
        gclo:leftProperties ( ?p1 ) ;
        gclo:classes ( ?c1 ) ;
        gclo:constrainingElement "universal quantification" ] .
    [   a gclo:ClassConstraint ;
        gclo:contextClass ?c1 ;
        gclo:classes ( ?i1 ?i2 ?i3 ) ;
        gclo:constrainingElement "union" ] .				
    ?subject a ?cc ; ?p1 ?i .
    FILTER ( ?i != ?i1 && ?i != ?i2 && ?i != ?i3 ) . }
\end{ex}

There can be no detailed difference in the automatic validation of a constraint when it is expressed by different constraint languages.
So, the validation results do not differ on the basis of multiple validation implementations, as there is only one validation implementation for each constraint independent from the used constraint language.

%\subsection{Not Allowed Values}

\section{Implementation}
\label{sec:implementation}
